---
title: "NZCPI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NZCPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(NZCPI)
```

Download latest CPI csv file (begins with 'consumers-price-index' ends with 'index-numbers.csv') from StatsNZ https://www.stats.govt.nz/large-datasets/csv-files-for-download/ and use as input in the 'create_cpi_df' function or use built in data set 'nz_cpi_data' (updated to June 2022) to view NZ's Consumer Price Index by quarter. 

You can use your own date range with 'start' and 'end' date arguments in all the functions or use the default start and end dates of the built in nz_cpi_data data set.  

Use the function 'plot_cpi' to display the CPI.

For detail on how the CPI has changed you can use the 'perc_change_cpi' function to see what the percentage change in CPI was between dates or use the 'apply_cpi' function on a single value or vector of values (such as column of a data frame) to adjust values for inflation and see how consumers purchasing power has changed over time.    
